### 2014-03-29 ms_probaweb.tex
A few refs and edits in the niche section

\newpage
%========================================================%
\section*{A probabilistic representation of ecological interactions networks}

@@

The niche concept is key to understand and predict species
distribution. Several attempts have been made to refresh it, but its main
usage still follows Hutchinson’s idea that species interactions restrict
the fundamental niche to a realized one, and ecologists haven't moved far past
the n-dimensional hypervolume formalism \parencite{Blonder2014}. Despite its
intuitive interpretation and translation into species distribution models,
the concept has been constantly criticized (Hardin, 1960; Peters, 1991;
Chase2003; Silvertown, 2004; Soberon, 2007) and several attempts have been
made to expand and reinforce it.

Part of the problem surrounding the definition of the niche has been
clarified with the distinction between Eltonian and Grinnellian definitions
(ChaseLeibold 2003). The Grinnellian dimension of the niche is the effect of
the environment on the demography of a species, while the Eltonian dimension
is the effect of a species on its environment \emph{sensu lato}.  % SNOB!
The Grinnellian niche is the most intuitive one to apply and is the conceptual
backbone of species distribution models. The Eltonian niche is well known
by community ecologists, but is trickier to turn into predictive models
\parencite{Devictor2010a}. Nonetheless, the development of the niche model
of food web structure (Williams2000) and its parameterization (Williams2010;
Gravel2013) made it more operational, although it has yet to be applied to
more than trophic interactions.

While it is straigthforward to represent statistically the hyper volume where a species
occurs, it is much more challenging to account for ecological interactions.
Chase and Leibold (2003) attempted this representation in their definition:
\textit{[The niche is] the joint description of the environmental conditions
that allow a species to satisfy its minimum requirements so that the birth rate
of a local population is equal or greater than its death rate along with the set
of per capita effects of that species on these environmental conditions.} They
represented the niche graphically with zero-net growth isoclines (the Grinnelian
niche) and impact vectors (the Eltonian niche). While this representation has
been very influential in community ecology at the local scale, it remains
unpracticable at the biogeographical one.
% Why (or more to the point, what do we care about having the same definition at different scales?)?
The absence of any mathematical
representation of the niche that could easily be fit to ecological data perhaps
explain why biogeographers are still struggling to develop species distribution
models taking into account ecological interactions.

>>>>

\newpage
%========================================================%
\section*{A probabilistic representation of ecological interactions networks in space}

@@

Correctly describing the ecological niche is key to understand and predict
species distribution. Despite several attempts to refresh the conceptual
basis of what ecological niches are, ecologists haven't moved far past the
"n-dimensional hypervolume" formalism \parencite{Holt2009} introduced by
\textcite{Hutchinson1961}. Despite its intuitive interpretation and translation
into species distribution models \parencite{Blonder2014}, the concept has been
constantly criticized (Hardin, 1960; Peters, 1991; Chase2003; Silvertown, 2004;
Soberon, 2007) and several attempts have been made to expand and reinforce it.

Part of the problem surrounding the definition of the niche has been
clarified with the distinction between Eltonian and Grinnellian definitions
\parencite{Chase2003}. The Grinnellian dimension of
the niche is the effect of the environment on the demography of a species,
while the Eltonian dimension is the effect of a species on its environment
\emph{sensu lato}.  % SNOB!
The Grinnellian niche is the most intuitive one
to apply and is the conceptual backbone of species distribution models. The
Eltonian niche is well known by community ecologists, but is trickier to
turn into predictive models \parencite{Devictor2010a}. Nonetheless, the
development of the niche model of food web structure (Williams2000) and
its parameterization (Williams2010; Gravel2013) made it more operational,
although it has yet to be applied to more than trophic interactions.

\textcite{Chase2003} attempted this representation in their definition:
\textit{[The niche is] the joint description of the environmental conditions
that allow a species to satisfy its minimum requirements so that the birth rate
of a local population is equal or greater than its death rate along with the
set of per capita effects of that species on these environmental conditions.}
Their representation merges zero-net growth figuring the Grinnelian niche
(when do the population persist) with impact vectors figuring the Eltonian
niche (what is the population impact). While this representation has been very
influential in community ecology at the local scale, it remains unpracticable
%%% WHY?
at the biogeographical one. The absence of any mathematical representation
of the niche that could easily be fit to ecological data perhaps explain why
biogeographers are still struggling to develop species distribution models
taking into account ecological interactions.
